Spike-and-Slab Generalized Additive Models and Scalable Algorithms for High-Dimensional Data

### 2022-03-18 Manuscript/05-Discussion.Rmd
Update Section Headers to Latex
revise discussion

# Discussion
\label{sec:concl}
In the paper, we described a novel generalized additive model using Bayesian hierarchical priors, particularly the proposed spike-and-slab spline prior for bi-level functional selection. Meanwhile, we introduced two optimization based algorithms for model fitting. The algorithms can be easily scale up to address high-dimensional data analysis in a computational efficient manner. Via simulations, we demonstrated that the proposed model provides as good, if not better, prediction performance compared to some state-of-the-art non-linear modeling devices.

The proposed model shares many commonality with an high-dimensional Bayesian GAM, SB-GAM [@Bai2021], independently developed around the same time of this work. Both frameworks emphasize computational efficiency by deploying group spike-and-slab lasso type of priors and optimization-based fast and scalable algorithms. Bai provided the theoretical proof for the consistency of variable selection using group spike-and-slab lasso prior. Nevertheless, SB-GAM fails to address the bi-level functional selection and the model inference. The proposed model provides solutions to these remaining questions while maintaining the same level of prediction accuracy. Moreover, the proposed model renders additional flexibility to model specification, allowing various choices of smoothing functions and degrees of freedom.

For translational science purpose, we implemented the proposed model in a R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM). To maximize the flexibility of smoothing function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smoothing functions are limited to the default ones. Ancillary functions are provided for model specification in high-dimensional settings, curve plotting and functional selection. In addition, in \texttt{BHAM}, we streamline the model fitting algorithms to support other popular Bayesian hierarchical prior for the smoothing functions, such as Student's T distribution, double exponential distribution, mixture normal and T distribution's. These priors could be helpful when the sparse assumption is weak or not necessary.,

<!-- the implementation of ssGAM-NM using EM-IRLS. Nevertheless, EM-IRLS are not limited to fitting ssGAM-NM. Other spike-and-slab prior can be fitted with EM-IRLS as well, e.g. mixture t-distribution, and mixture double exponential distributions. @Yi2012 demonstrated that t prior and double exponential prior can be seen hierarchical normal priors, -->
<!-- $$ -->
<!-- \begin{aligned} -->
<!-- \beta_i \sim N(0, \tau^2_i),\\ -->
<!-- \tau^2_i \sim Inv-\chi^2(v, s^2_j) \text{ or } \tau^2_i \sim Gamma(1, s^2/2) -->
<!-- \end{aligned} -->
<!-- $$ -->
<!-- with $\tau^2$ integrated out analytically. Hence, the EM-IRLS can be easily applied to those two additional mixture priors. One of the advantage of using mixture $t$ prior over mixture normal prior is that $t$ distributions with low degree of freedom, e.g. $v=1$, have heavier tail than the normal distribution, which is also a special case of $t$ distribution when $v=\infty$. The heavier tails help to obtain posteriors that concentrated around the true nonzero coefficients at optimal rates. [@Rockova2014a] -->

There are some improvements possible for the proposed models. First of all, the proposed model achieves a bi-level selection via the two-part spike-and-slab spline prior. Nevertheless, this set-up could result in a situation that is not theoretically sound: the non-linear component is selected, but the linear component is not. We currently address it analytically by including the linear component in the model when non-linear component is selected. Another possible solution is to impose a dependent structure of $\gamma_{j}^\tp$ on $\gamma_{j^{0}}$, i.e. $\gamma_j^\tp|\gamma_{j}^{0}, \theta_j$. Secondly, the computational time for fitting a BHAM model with EM-IWLS algorithms can be improved. Current implementation of the EM-IWLS algorithm jointly updates all coefficients in each iteration, which requires a lot computation resources. This jointly updating procedure can be enhanced by adapting a backfitting step [@Hastie1987] where each smoothing function are updated individually. Thirdly, when using the proposed model in real data analysis where a screening procedure is implemented before joint predictive modeling. we recommend to include the screening procedure in the cross-validation during model selection. [@Friedman2017] For the sake of model comparison, we fit the models using the same pool of predictors in Section \ref{sec:ECB}.

@@

  <!-- Such prior set-ups ignore the null space of the smooth function, and entirely removes the variable out of the function. The "all-in-all-out" assumption makes differentiate linear and non-linear effect harder. In contrasts, our prior set up provides a more flexible solution, as the penalty for smoothness is locally adapative for the bases and hence retain the the null space if necessary.  -->

<!-- The "all-in-all-out" set-up was criticized for its inability to answering an important research question, if the effects are non-linear or not. On the other hand, the model can be much more flexible by allowing each dimension of the penalized space to have an independent indicator, i.e. $\gamma^\tp_{jk}$ of $\beta^\tp_{jk}$ for $k = 1\dots, K_j$. This prior set-up generalized to the group spike-and-slab. The additional flexibility doesn't not necessarily grand the model neither theoretical justification nor analytic simplicity. The spline penalty assumes a common smooth shrinkage $\lambda_j$ applied to all coefficients of the penalized space, which contradicts to the model setup. Besides, having multiple indicator variables complicates the decision of if non-linear effects are necessary. -->

>>>>

\section{Discussion}
\label{sec:concl}
\bg{In the paper, we described a novel high-dimensional generalized additive model with Bayesian hierarchical prior for the purpose of predictive modelling. In particular, we introduce a two-part spike-and-slab lasso prior for re-parameterized smoothing function and derive a scalable EM-CD algorithm for model fitting. The proposed model provides a new angle to address the excess shrinkage of smoothing functions that is commonly vulnerable to previous regularized high-dimensional GAMs, and hence improves the predictive performance. Th EM-CD algorithm, extended from previous spike-and-slab regression models, provides a computationally efficient alternative to the computational prohibitive MCMC algorithms for model fitting. Meanwhile, the two-part prior motivates the bi-level selection of functional forms of the predictors, selection of linear and nonlinear component. In the simulation and real-data analysis, the proposed model demonstrates improvement in prediction and computational advantage when compared to some state-of-the-art models. When serving the purpose of variable selection, trade-offs exist among methods of comparison.} We implement the proposed model in an open-source R package \texttt{BHAM}, deposited at [https://github.com/boyiguo1/BHAM](https://github.com/boyiguo1/BHAM). To maximize the flexibility of smoothing function specification, we deploy the same programming grammar as in the state-of-the-art package \texttt{mgcv}, in contrast to previous tools where smoothing functions are limited to the default ones. Ancillary functions are provided for model specification in high-dimensional settings, curve plotting and functional selection.

The proposed model shares many commonality with an high-dimensional Bayesian GAM, SB-GAM [@Bai2021], independently developed around the same time of this work. Both frameworks emphasize computational efficiency by deploying group spike-and-slab lasso type of priors and optimization-based scalable algorithms. Nevertheless, Bai provided the theoretical proof for the consistency of variable selection using group spike-and-slab lasso prior. \bg{The proposed model focuses on improving prediction performance for high-dimensional GAM, with the capability of bi-level selection. Moreover, the proposed model can be generalized to other family of priors or other family of smoothing functions if desired. Not focused in this manuscript, the generalization is described in the supporting materials.}

\bg{During designing and analyzing the simulation study, we made couple interesting observations. First of all, variable selection is a delicate topic in the context predictive modelling. When prediction performance is used to tune a model, the model could possibly include noise variable in models, for example Lasso. [TODO: add Denilla paper] In addition, bi-level selection is a more complex problem than variable selection. The complexity reflects on the validity of effect hierarchy assumption. While most functional forms follow effect hierarchy that linear components exists in the nonlinear function, there are functions that don't follow it, e.g. $x^2$. The proposed prior and spikeSlabGAM employ different structure: the proposed prior imposes effect hierarchy while spikeSlabGAM treats the selection of linear and nonlinear components independent. The different prior setups lead to trade-offs for the purpose of bi-level selection. We recommend to use more judgement in bi-level selection, either relying on heuristic knowledge to choose appropriate prior or exploring multiple models when heuristic knowledge doesn't exist. Secondly, we find the performance of the proposed model are more sensitive to the granularity of $s_0$ sequence in the high-dimensional settings than in the lower dimension settings. Even though the current default sequence of $s_0$ can result in reasonable performance shown in the simulation studies, we recommend to fine-tune the model with granular sequence of $s_0$ for performance improvement.}

<!-- There are some improvements possible for the proposed models. First of all, the proposed model achieves a bi-level selection via the two-part spike-and-slab spline prior. Nevertheless, this set-up could result in a situation that is not theoretically sound: the non-linear component is selected, but the linear component is not. We currently address it analytically by including the linear component in the model when non-linear component is selected. Another possible solution is to impose a dependent structure of $\gamma_{j}^\tp$ on $\gamma_{j^{0}}$, i.e. $\gamma_j^\tp|\gamma_{j}^{0}, \theta_j$. Secondly, the computational time for fitting a BHAM model with EM-IWLS algorithms can be improved. Current implementation of the EM-IWLS algorithm jointly updates all coefficients in each iteration, which requires a lot computation resources. This jointly updating procedure can be enhanced by adapting a backfitting step [@Hastie1987] where each smoothing function are updated individually. Thirdly, when using the proposed model in real data analysis where a screening procedure is implemented before joint predictive modeling. we recommend to include the screening procedure in the cross-validation during model selection. [@Friedman2017] For the sake of model comparison, we fit the models using the same pool of predictors in Section \ref{sec:ECB}. -->

@@
